Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-07-08 paper/main.tex
minor update

The rule {\sc L:Call} accounts for the execution cost of both
$M_f$ for function calls and $M_r$ for \code{return} statements.  \iffull{This
justifies that the {\sc L:Return} rule does not account for resource
consumption.}{}  The pre- and postcondition $P_f$ and $Q_f$ are taken
from the function context $\Delta$.  The assertions in the context are
parametric with respect to both the function arguments' value and the
return value. This allows to specify a bound for a function whose
resource consumption depends on input parameters.  The argument
parameters are instantiated by the call rule using the result of the
evaluation of the argument variables in the current state.
\iffull{Recall that we require that $P_f$ and $Q_f$ do not depend on
  the local variables on the call stack, that is, $\forall z \, \vec v
  \, \theta \, \theta' \, \gamma . \, P_f \, z \, \vec v \,
  (\theta,\gamma) = P_f \, z \, \vec v \, (\theta',\gamma)$.}{} To
transfer potential that depends on local variables of the callee from
the precondition $P$ to the postcondition $Q$, we use an assertion $A
: \Assn$ that is independent of global variables, that is, $\forall
\theta \, \gamma \, \gamma' . \, A (\theta,\gamma) =
A(\theta,\gamma')$.  It is still possible to express relations between
global and local variables using logical variables\iffull{ (see the
following paragraph for details).}{.}

@@

To find derivations in the quantitative Hoare logic automatically, we
have to focus on bounds that have a certain shape.  The general form
of \emph{potential functions} (or assertions) that we consider is
$$
  \Phi(\state) = q_0 + \sum_{x, y \in \dom\state \land x \neq y}
    q_{(x, y)} \cdot |\inter {\state(x)} {\state(y)}| \, .
$$
%
Here $\state : (\text{Locals} \to \Z) \times (\text{Globals} \to \Z)$
is again a simplified program state as introduced in \pref{sec:sem},
$|\inter a b| = max(0,b-a)$, and $q_i \in \Qplusz$.  To simplify the
references to the linear coefficients $q_i$, we introduce an \emph{index
  set} $I$.  This set is defined to be $\{0\} \cup \{(x, y) \mid x, y
\in \Var \land x \neq y \}$.  Each index $i$ corresponds to a \emph{base function}
$f_i$ in the potential function: $0$ corresponds to the constant
function $\state \mapsto 1$, and $(x,y)$ corresponds to $\state \mapsto
|\inter {\state(x)} {\state(y)}|$.  Using these notations we can
rewrite the above equality as\iffull{
$$
  \Phi = \sum_{i \in I} q_i f_i.
$$}
{ $\Phi = \sum_{i \in I} q_i f_i$.}  We often write $xy$ to denote the
index $(x,y)$.  The family $(f_i)_{i\in I}$ is actually a basis (in
the linear-algebra sense) for potential functions.  This allows us to
uniquely represent any linear potential function
$\Phi$ as a \emph{quantitative annotation} $Q = (q_i)_{i \in I}$, that
is, a family of \ifshort{non-negative}{} rational numbers where only
a finite number of elements are not zero. \iffull{ For soundness, we
  require that potential functions always give non-negative results.
  This is acheived by restricting coefficients in a quantitative
  annotation to be always non-negative.}{}

>>>>

The rule {\sc L:Call} accounts for the execution cost of both $M_f$
for function calls and $M_r$ for \code{return} statements.
\iffull{This justifies that the {\sc L:Return} rule does not account
  for resource consumption.}{} The pre- and postcondition $P_f$ and
$Q_f$ are taken from the function context $\Delta$.  The assertions in
the context are parametric with respect to both the values of the
function arguments and the return value. This allows us to specify a
bound for a function whose resource consumption depends on its
arguments.  The arguments are instantiated by the call rule
using the result of the evaluation of the argument variables in the
current state.  \iffull{Recall that we require that $P_f$ and $Q_f$ do
  not depend on the local variables on the call stack, that is,
  $\forall z \, \vec v \, \theta \, \theta' \, \gamma . \, P_f \, z \,
  \vec v \, (\theta,\gamma) = P_f \, z \, \vec v \,
  (\theta',\gamma)$.}{} To transfer potential that depends on local
variables of the callee from the precondition $P$ to the postcondition
$Q$, we use an assertion $A : \Assn$ that is independent of global
variables, that is, $\forall \theta \, \gamma \, \gamma' . \, A
(\theta,\gamma) = A(\theta,\gamma')$.  It is still possible to express
relations between global and local variables using logical
variables\iffull{ (see the following paragraph for details).}{.}

@@

To find derivations in the quantitative Hoare logic automatically, we
have to focus on bounds that have a certain shape.  The general form
of \emph{potential functions} (or assertions) that we consider is
$$
  \Phi(\state) = q_0 + \sum_{x, y \in \dom\state \land x \neq y}
    q_{(x, y)} \cdot |\inter {\state(x)} {\state(y)}| \, .
$$
%
Here $\state : (\text{Locals} \to \Z) \times (\text{Globals} \to \Z)$
is again a simplified program state as introduced in \pref{sec:sem},
$|\inter a b| = max(0,b-a)$, and $q_i \in \Qplusz$.  To simplify the
references to the linear coefficients $q_i$, we introduce an \emph{index
  set} $I$.  This set is defined to be $\{0\} \cup \{(x, y) \mid x, y
\in \Var \land x \neq y \}$.  Each index $i$ corresponds to a \emph{base function}
$f_i$ in the potential function: $0$ corresponds to the constant
function $\state \mapsto 1$, and $(x,y)$ corresponds to $\state \mapsto
|\inter {\state(x)} {\state(y)}|$.  Using these notations we can
rewrite the above equality as\iffull{
$$
  \Phi = \sum_{i \in I} q_i f_i.
$$}
{ $\Phi = \sum_{i \in I} q_i f_i$.}  We often write $xy$ to denote the
index $(x,y)$.  \iffull{The family $(f_i)_{i\in I}$ is actually a basis (in
the linear-algebra sense) for potential functions.}{}  This allows us to
uniquely represent any linear potential function
$\Phi$ as a \emph{quantitative annotation} $Q = (q_i)_{i \in I}$, that
is, a family of \ifshort{non-negative}{} rational numbers where only
a finite number of elements are not zero. \iffull{ For soundness, we
  require that potential functions always give non-negative results.
  This is acheived by restricting coefficients in a quantitative
  annotation to be always non-negative.}{}
